Exploring probability distributions for bivariate temporal granularities

### 2019-09-23 main.Rmd
cricket examples updated

# Hourly usage across days of the week
sm_cust50 %>%
  granplot("day_week", 
           "hour_day",
           "general_supply_kwh", 
           plot_type = "quantile", 
          quantile_prob = c(0.25, 0.5, 0.75),
           response = "general_supply_kwh") +
  ggtitle("") +
  ggtitle("Quantile plot across hour-of-day and day-of-week") +
  scale_x_discrete(breaks = seq(0, 23, 3)) +
  scale_color_brewer(type = "div",
                     palette = "Dark2", direction = -1)
#observation: on saturday and sunday morning peaks are higher and little late (in absolte time) than other weekdays

# Across days of the week by months
sm_cust50 %>%
  granplot("month_year",
    "day_week",
    response = "general_supply_kwh", plot_type = "quantile", 
          quantile_prob = c(0.25, 0.5, 0.75)) + ggtitle("") +
  ggtitle("Quantile plot across day-of-week by month-of-year")

@@

@@

```{r runs_per-over}
cricket_per_over <- cricketdata %>% 
  group_by(match_id, batting_team, bowling_team,  inning, over) %>%
  summarise(runs_per_over = sum(total_runs))

cricket_tsibble <- cricket_per_over %>% 
  ungroup() %>% 
  mutate(data_index = row_number()) %>% 
  as_tsibble(index = data_index)
```

First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas_over_inning}. The distribution of runs per over has not significantly changed from 2008 to 2016. There is no clear trend that we can observe for any over of the innings for any seasons. So we can work with any season and

```{r seas_over_inning}
 cricket_tsibble_all %>% granplot("over_inning", "season",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

  )  + 
    theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_line( size=.1, color="black" ) 
    )

@@

  cricket_tsibble %>% granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
    theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_line( size=.1, color="black" ) 
    )

  cricket_tsibble %>%
    filter(batting_team == "Chennai Super Kings") %>% 
    granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
  )

    cricket_tsibble %>%
    filter(batting_team == "Kolkata Knight Riders") %>% 
    granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
  )

@@

# all season per over runs
cricket_per_over_all <- cricket_all_seas %>% 
  group_by(season, match_id, batting_team,inning, over) %>%
  summarise(runs_per_over = sum(total_runs))

# all season per over runs tsibble
cricket_tsibble_all <- cricket_per_over_all %>% 
  ungroup() %>% 
  mutate(data_index = row_number()) %>% 
  as_tsibble(index = data_index)

  cricket_tsibble_all %>% granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

  )

  cricket_tsibble_all %>% granplot("inning_match", "season",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

  )

  cricket_tsibble_all %>% granplot("season", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

  )  +

  cricket_tsibble %>%
    filter(batting_team == "Chennai Super Kings") %>% 
    granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
  )

    cricket_tsibble %>%
    filter(batting_team == "Kolkata Knight Riders") %>% 
    granplot("inning_match", "over_inning",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
  )

@@

```{r read_data_all_seas, echo=FALSE}

load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda")

cricket_tsibble_all <- cricket_all_seas %>%
  filter(season %in% c(2012:2015)) %>% 
  mutate(data_index = row_number()) %>%
  as_tsibble(index = data_index)

# cricket_tsibble %>% dynamic_create_gran(
#   "ball_over",
#   hierarchy_model
# )

#over_inning    ball_over 
cricket_tsibble_all %>% granplot("over_inning", "ball_over",
  hierarchy_model,
  response = "total_runs",
  plot_type = "quantile",
  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

)

#inning_match   ball_over

cricket_tsibble_all %>% granplot("inning_match", "ball_over",
  hierarchy_model,
  response = "total_runs",
   plot_type = "quantile",
  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)
)

# inning_match   ball_inning 
#interesting plot
cricket_tsibble_all %>% granplot("inning_match", "ball_inning",
  hierarchy_model,
  response = "total_runs",
   plot_type = "quantile",
  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+
  scale_x_discrete(breaks = seq(1,120,5))

# ball_over      over_inning

cricket_tsibble_all %>% granplot("ball_over", "over_inning",
  hierarchy_model,
  response = "total_runs",
  plot_type = "quantile",
  quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

# inning_match   over_inning

cricket_tsibble_all %>% granplot("inning_match", "over_inning",
  hierarchy_model,
  response = "total_runs"
)

# ball_over      over_match

cricket_tsibble_all %>% granplot("ball_over", "over_match",
  hierarchy_model,
  response = "total_runs"
)

# ball_over      inning_match

cricket_tsibble_all %>% granplot("ball_over", "inning_match",
  hierarchy_model,
  response = "total_runs"
) + scale_color_brewer(type = "div", palette = "Dark2")

cricket_tsibble_all %>%
  granplot(
    gran1 = "inning_match",
    gran2 = "over_inning",
    hierarchy_model,
    response = "total_runs",
    plot_type = "quantile"
  )
```

>>>>

# Hourly usage across days of the week
sm_cust50 %>%
  granplot("day_week", 
           "hour_day",
           "general_supply_kwh", 
           plot_type = "quantile", 
          quantile_prob = c(0.25, 0.5, 0.75),
           response = "general_supply_kwh", overlay = FALSE) +
  ggtitle("") +
  ggtitle("Quantile plot across hour-of-day and day-of-week") +
  scale_x_discrete(breaks = seq(0, 23, 3)) +
  scale_color_brewer(type = "div",
                     palette = "Dark2", direction = -1)
#observation: on saturday and sunday morning peaks are higher and little late (in absolte time) than other weekdays

# Across days of the week by months
sm_cust50 %>%
  granplot("month_year",
    "day_week",
    response = "general_supply_kwh", plot_type = "quantile", 
          quantile_prob = c(0.25, 0.5, 0.75), overlay = FALSE) + ggtitle("") +
  ggtitle("Quantile plot across day-of-week by month-of-year")

@@

Is the outcome of the game dependent on who bats first? Do some teams have more chance to win if they are batting in the first innings?

@@

cricket <-  read_csv("data-raw/deliveries_all.csv")
matches <- read_csv("data-raw/matches_all.csv")

cricket_season <- cricket %>% left_join(matches, by = c("match_id" ="id"))

cricket_per_over <- cricket_season %>% 
  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
  summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs))

cricket_tsibble_all <- cricket_per_over %>% 
  ungroup() %>% 
  mutate(data_index = row_number()) %>% 
  as_tsibble(index = data_index)
```

First, we look at the distribution of runs per over across over of the innings and seasons in \autoref{fig:seas_over_inning}. The distribution of runs per over has not significantly changed from 2008 to 2016. There is no clear trend that we can observe for any over of the innings for any seasons. So we can work with any subsets of seasons and try to see how the strategies of the winning teams differ across seasons or how in a particular season the strategy was different for the winning team and the ones who did not qualify for the playoffs.

```{r seas_over_inning}
 cricket_tsibble_all %>% granplot("over", "season",
                               hierarchy_model,
                               response = "runs_per_over",
                               plot_type = "quantile",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)

  )  + 
    theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_blank() 
    ) + scale_x_discrete(breaks = seq(2008, 2016, 3))

@@

Mumbai Indians(MI) won against Chennai Super kings(CSK) in 2013 and 2015. So it would be interesting to see the difference in the strategies between these two teams tgroughout all matches in the two seasons. In \autoref{fig:2013_15_MI}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match. It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings. CSK has been more uniform in their approach maintaining consistent scores per over, whereas MI's scores increase as they proceed with the game nothwithstanding if they bat first or second. CSK's run rate increase as it reached the last over when it bats in the 2nd innings, but doesn't increase much when it bats first.

<!-- ```{r 2013_15_MI} -->
<!--  cricket_tsibble_all %>%  -->
<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
<!--            granplot("inning", "over", -->
<!--                                hierarchy_model, -->
<!--                                response = "run_rate", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
<!--     ) -->
<!--  cricket_tsibble_all %>%  -->
<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
<!--            granplot("inning", "over", -->
<!--                                hierarchy_model, -->
<!--                                response = "run_rate", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
<!--     ) -->

```{r lv_MI_CSK}
# fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Letter value plot of run rate across overs of an inning and innings of the match for (a) Mumbai Indians and (b) Chennai Super Kings.", fig.show = 'hold'

 cricket_tsibble_all %>% 
  filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>% 
           granplot("inning", "over",
                               hierarchy_model,
                               response = "run_rate",
                               plot_type = "lv",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("")

 cricket_tsibble_all %>% 
  filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>% 
           granplot("inning", "over",
                               hierarchy_model,
                               response = "run_rate",
                               plot_type = "lv",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("")

@@

```{r MI_CSK_violin}

matches_select <- matches %>% select(id, toss_winner, winner)

cricket_winner <- cricket_tsibble_all %>% 
  left_join(matches_select, by = c("match_id" = "id"))

MI_winner <- cricket_winner %>% 
             filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians") %>% 
             mutate(win = if_else(winner == "Mumbai Indians", 1, 0))

CSK_winner <- cricket_winner %>% 
             filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings") %>% 
             mutate(win = if_else(winner == "Chennai Super Kings", 1, 0))

 MI_winner %>% 
            granplot( "win","inning",
                                hierarchy_model,
                                response = "run_rate",
                                plot_type = "violin",
                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
                     overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)

 CSK_winner %>% filter(inning %in% c(1,2)) %>% 
           granplot( "win","inning",
                               hierarchy_model,
                               response = "run_rate",
                               plot_type = "violin",
                               quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
                    overlay = FALSE) + ggtitle("") + geom_boxplot(width = 0.1)

<!-- For MI, cases when they lost, distribution of run rates not very different depending on if they bat first or second. However, cases when they won, distribution of run rate is very skewed when they bat in the 2nd innings implying winning by short margins.  -->

For CSK, irrepective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing. Run rate can go up to 6 runs per over depending on the situation. For MI, this is not true. When they won, their run rates in the 2nd innings went up to 6 and led to their win, but stayed within 4 whenever they lost. This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings.

<!--   cricket_tsibble %>% granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
<!--     ) -->

<!--   cricket_tsibble %>% -->
<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
<!--     granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!--     cricket_tsibble %>% -->
<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
<!--     granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!-- ```{r runs_per-over} -->

<!-- # all season per over runs -->
<!-- cricket_per_over_all <- cricket_all_seas %>%  -->
<!--   group_by(season, match_id, batting_team,inning, over) %>% -->
<!--   summarise(runs_per_over = sum(total_runs))  -->

<!-- # all season per over runs tsibble -->
<!-- cricket_tsibble_all <- cricket_per_over_all %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(data_index = row_number()) %>%  -->
<!--   as_tsibble(index = data_index) -->

<!--   cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   cricket_tsibble_all %>% granplot("inning_match", "season", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   cricket_tsibble_all %>% granplot("season", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  +  -->

<!--   cricket_tsibble %>% -->
<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
<!--     granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!--     cricket_tsibble %>% -->
<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
<!--     granplot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

@@

<!-- ```{r read_data_all_seas, echo=FALSE} -->

<!-- load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda") -->

<!-- cricket_tsibble_all <- cricket_all_seas %>% -->
<!--   filter(season %in% c(2012:2015)) %>%  -->
<!--   mutate(data_index = row_number()) %>% -->
<!--   as_tsibble(index = data_index) -->

<!-- #over_inning    ball_over  -->
<!-- cricket_tsibble_all %>% granplot("over_inning", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!-- ) -->

<!-- # inning_match   ball_inning  -->
<!-- #interesting plot -->
<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
<!--   scale_x_discrete(breaks = seq(1,120,5)) -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->

<!-- cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "over_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "inning_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->

<!-- cricket_tsibble_all %>% -->
<!--   granplot( -->
<!--     gran1 = "inning_match", -->
<!--     gran2 = "over_inning", -->
<!--     hierarchy_model, -->
<!--     response = "total_runs", -->
<!--     plot_type = "quantile" -->
<!--   ) -->
<!-- ``` -->
